Mapping

### 2024-04-08 episodes/quality-control.Rmd
Fixed a few errors in QC Adapted the mapping part

We can now see that the quality of an `N` (`'`) is 6, which corresponds to a over 25% error probability (\$10\^{-\frac{6}{10}} \cong 0.25 \$).

> ### Quality Encodings Vary
>
> Although we've used a particular quality encoding system to demonstrate interpretation of read quality, different sequencing machines use different encoding systems. This means that, depending on which sequencer you use to generate your data, a `#` may not be an indicator of a poor quality base call.

@@

``` output
...
ERR029206_1_fastqc.html     ERR029206_1.fastq.gz_trim.fastq  ERR029206_2_fastqc.zip
ERR029206_1_fastqc.zip	    ERR029206_1_trim.fastq		       ERR029206_2.fastq.gz
ERR029206_1.fastq.gz        ERR029206_2_fastqc.html	         ERR029206_2.fastq.gz_trim.fastq
```

@@

::: challenge
### Challenge

@@

@@

Coverage = #bases (forward + reverse) / genome size In this case: 125.18 = (287211920 + 265048852) / 4411532
:::
:::

>>>>

We can now see that the quality of an `N` (`'`) is 6, which corresponds to a over 25% error probability:

$10^{-\frac{6}{10}} \cong 0.25 $

### Quality Encodings Vary

Although we've used a particular quality encoding system to demonstrate interpretation of read quality, different sequencing machines use different encoding systems. This means that, depending on which sequencer you use to generate your data, a `#` may not be an indicator of a poor quality base call.

@@

@@

::: challenge

### Challenge

@@

::: solution

### Solution

@@

Coverage = #bases (forward + reverse) / genome size In this case:

$125.18 = \dfrac{(287211920 + 265048852)}{4411532}$

:::
:::

### 2024-04-08 episodes/snps-phylogeny.Rmd
Fixed a few errors in QC Adapted the mapping part

```bash
cd ~/molepi/data
mv trimmed_fastq trimmed_fastq_partial
wget https://uppsala.box.com/s/s1zpt0w84iq4o42qhgho5vrz9qnpxouk trimmed_fastq.tar.gz
tar xvzf trimmed_fastq.tar.gz
ls trimmed_fastq
```

@@

```bash
$ cd ~/molepi/results
$ mkdir snps
```

We need to activate a separate environment to run snippy:

```bash
$ mamba activate snippy
```

@@

```bash
$ cd ../data/trimmed_fastq/
$ #for sample in ERR026473 ERR026474 ERR026478 ERR026481 ERR026482 ERR029206 ERR029207
$ for sample in ERR029206
do
snippy --ram 1 --outdir ../../results/snps/"${sample}" --ref ../GCF_000195955.2_ASM19595v2_genomic.fna --R1 "${sample}"_1.fastq.gz_trim.fastq --R2 "${sample}"_2.fastq.gz_trim.fastq
done
```

Here, we provide snippy with an output folder (`--outdir`), the location of the reference genome (`--ref`), and the trimmed read files for each end of the pair (`--R1` and `--R2`). We also indicate that snippy should use 1 Gb of memory (`--ram 1`, very specific issue to this VM).

```bash
$ head -n10 ~/molepi/results/snps/ERR029206/snps.tab 
```

@@

Now download a [reduced version of the result of snippy](files/snps.tar.gz) on the other samples, save it in the `~/molepi/results/` folder. We will also rename the existing `snps` folder to save it.

```bash
cd ~/molepi/results/
mv snps snps_partial
mkdir snps
cd snps
tar xvzf ../snps.tar.gz
ls
```

@@

```bash
$ cat ~/molepi/results/snps/ERR029207/snps.txt
```

@@

```bash
$ cd ~/molepi/results/snps/
$ snippy-core --ref=../../data/GCF_000195955.2_ASM19595v2_genomic.fna ERR026473 ERR026474 ERR026478 ERR026481 ERR026482 ERR029206 ERR029207
```

@@

Let's deactivate the specific snippy `env` before we move on:

```bash
$ mamba deactivate snippy
```

@@

```bash
$ iqtree -s core.aln -m MFP+ASC
```

@@

```bash
$ cd ~/molepi/results
$ mkdir tree
$ mv snps/core.aln.* tree
```

@@

```bash
$ cd ~/molepi/results/
$ cd tree
$ mv core.aln.treefile core.aln.newick
$ core.aln.newick
```

@@

```bash
$ cat core.aln.iqtree
```

>>>>

```bash
interactive -A uppmax2024-2-10 -M snowy -t 04:00:00
cd /proj/g2020004/nobackup/3MK013/<username>/molepi/data
mv trimmed_fastq trimmed_fastq_partial
tar xvzf ../../../data/trimmed_fastq.tar.gz
ls trimmed_fastq
```

@@

```bash
interactive -A uppmax2024-2-10 -M snowy -t 04:00:00
cd /proj/g2020004/nobackup/3MK013/<username>/molepi/data
mkdir snps
```

We need to load snippy:

```bash
module load snippy
```

@@

```bash
cd ../data/trimmed_fastq/
#for sample in ERR026473 ERR026474 ERR026478 ERR026481 ERR026482 ERR029206 ERR029207
for sample in ERR029206
do
snippy --ram 8 --outdir ../../results/snps/"${sample}" --ref ../GCF_000195955.2_ASM19595v2_genomic.fna --R1 "${sample}"_1.fastq.gz_trim.fastq --R2 "${sample}"_2.fastq.gz_trim.fastq
done
```

Here, we provide snippy with an output folder (`--outdir`), the location of the reference genome (`--ref`), and the trimmed read files for each end of the pair (`--R1` and `--R2`). We also indicate that snippy should use 8 Gb of memory (`--ram 8`).

```bash
head -n10 ~/molepi/results/snps/ERR029206/snps.tab 
```

@@

Now copy a reduced version of the results on the other samples, save it in the `results` folder. We will also rename the existing `snps` folder to save it.

```bash
cd /proj/g2020004/nobackup/3MK013/<username>/molepi/results
mv snps snps_partial
mkdir snps
cd snps
tar xvzf ../../../../data/snps.tar.gz
ls
```

@@

```bash
cat ~/molepi/results/snps/ERR029207/snps.txt
```

@@

```bash
cd /proj/g2020004/nobackup/3MK013/<username>/molepi/results
snippy-core --ref=../../data/GCF_000195955.2_ASM19595v2_genomic.fna ERR026473 ERR026474 ERR026478 ERR026481 ERR026482 ERR029206 ERR029207
```

@@

@@

```bash
module load iqtree
iqtree2 -s core.aln -m MFP+ASC
```

@@

```bash
cd /proj/g2020004/nobackup/3MK013/<username>/molepi/results
mkdir tree
mv snps/core.aln.* tree
```

@@

```bash
cd /proj/g2020004/nobackup/3MK013/<username>/molepi/results
cd tree
mv core.aln.treefile core.aln.newick 
cat core.aln.newick
```

@@

```bash
cat core.aln.iqtree
```
